Cloud Supercomputing

### 2019-05-02 writing/paper/paper.tex
More work on paper.

\usepackage{caption}
\usepackage{xparse} % for NewDocumentCommand
\usepackage{etoolbox} % for notblank (brackets only when argument)
\usepackage{xstring} % for \IfSubStr
\usepackage{xpatch}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools} % for \mathclap
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage{varioref}
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage[separate-uncertainty]{siunitx}
\usepackage{hyperref}
\usepackage[noabbrev]{cleveref}
\newcommand{\creflastconjunction}{, and\nobreakspace} % use Oxford comma
\usepackage{todonotes}
\usepackage{multimedia}
\graphicspath{{../figures/}}
\let\boundary\undefined%
\crefformat{equation}{\eqA{}eq.~\eqB #2#1#3)}
% TODO: Fix for more than two equations!
\crefmultiformat{equation}{eqs.~\eqMultiA#2#1#3\eqMultiB}%
{ and \eqMultiA#2#1#3\eqMultiB}{, \eqMultiA#2#1#3\eqMultiB}{ and~\eqMultiA#2#1#3\eqMultiB}
\newcommand{\eqA}{}
\newcommand{\eqB}{(}
\newcommand{\eqMultiA}{(}
\newcommand{\eqMultiB}{)}
\DeclareRobustCommand{\pcrefSingle}[1]{%
\begingroup%
  \renewcommand{\eqA}{(}\renewcommand{\eqB}{}%
\cref{#1}%
\endgroup%
}
\DeclareRobustCommand{\pcrefMulti}[1]{%
\begingroup%
    \renewcommand{\eqMultiA}{}\renewcommand{\eqMultiB}{}%
    (\cref{#1})%
\endgroup%
}
\DeclareRobustCommand{\pcref}[1]{%
\IfSubStr{#1}{,}{\pcrefMulti{#1}}{\pcrefSingle{#1}}%
}

@@

We follow a different approach in this paper and use the \aderdg{} method of~\cite{dumbser2008unified} to simulate various computational fluid dynamics (\textsc{cfd}) scenarios with a focus on flows with stratified flows.
Due to the viscous components of the Navier-Stokes equations, it is not straightforward to apply the \dg{} method.
For high-order \dg{} methods, viscosity not only models the physical reality of the problem but also helps us by smoothing out oscillations and discontinuities, thus stabilizing the simulation.
To include viscosity, we use the numerical flux for the compressible Navier-Stokes equations of \citeauthor{gassner2008discontinuous}~\cite{gassner2008discontinuous}.
This flux has already been applied to the \aderdg{} method in~\cite{dumbser2010arbitrary}.
In contrast to this paper, we focus on the simulation of complex flows with a gravitational source term and a realistic background atmosphere.
Additionally, we use adaptive mesh refinement (\textsc{amr}) to increase the spatial resolution in areas of interest.
This has been shown to work well for the simulation of cloud dynamics~\cite{muller2010adaptive}.
%\todo{Einordnung, Kontext zu giraldo, etc$\ldots$}
%Insert references to Giraldo SE/DG bubbles~\cite{giraldo2008study}, 3D bubbles~\cite{giraldo2013implicit,kelly2012continuous}, bubble amr (müller~\cite{muller2010adaptive}).

@@

We make the following contributions:
\begin{itemize}%
\item We extend the \exahypeengine{} to allow viscous terms following the aforementioned numerical scheme.
\item We use this to provide an implementation of the compressible Navier-Stokes equation.
  In addition, we tailor the equation set to stratified flows with gravitational source term.
  We want to emphasize here that we use a standard formulation of the Navier-Stokes equations as seen in the field of computational fluid mechanics and only use small modifcations of the governing equations, in contrast to a equation set that is tailored exactly to the application area.
\item We present a general \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
  Furthermore, we show how to utilize this criterion for stratified flows.
\item We evaluate our implementation with standard \textsc{cfd} scenarios and atmospheric flows and inspect the effectiveness of our proposed \amr{}-criterion.
  We thus inspect, whether our proposed general implementation can achieve results that are competitive with the state-of-the-art models that rely on heavily specified equations and numerics.
\end{itemize}

\section{Equation Set}
\newcommand{\diffCoeff}{\varepsilon}%
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure)
  \end{pmatrix}
}%
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
    0\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}
   \end{pmatrix}
}%

\begin{equation}
  \label{eq:ns-short}
 \pdv{}{t} \bm{Q} + \divergence{\bm{F}(\bm{Q}, \gradient{\bm{Q}})} = \bm{S}(\bm{Q}) 
\end{equation}
The compressible Navier-Stokes equations in the conservative form are defined as% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
with vector of conserved quantities $\Q$, flux $\flux(\Q, \gradQ)$ and source $\source(\Q)$.
Note that the flux can be split into a hyperbolic part $\hyperFlux(\Q)$,
which is identical to the flux of the Euler equations,
and a viscous part $\viscFlux(\Q, \gradQ)$.
The conserved quantities
\(\Q\)
% \begin{equation} 
%   \label{eq:conserved-variables}
%  \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
% \end{equation}
are the density $\Qrho$, the two or three-dimensional momentum $\Qj$ and the energy density $\QE$.
The rows of \cref{eq:equation-set} are the conservation of mass, the conservation of momentum and the conservation of energy.

@@

Many realistic atmospheric flows can be described by a perturbation over a background state that is in hydrostatic equilibrium\todo{Mention source}
\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
i.e.\ a state, where the pressure gradient is exactly in balance with the gravitational source term $\source[\Qj] = - \bm{k} \Qrho g$.
The vector $\bm{k}$ is the unit vector pointing in $z$-direction.
The momentum equation is dominated by the background flow in this case.
Because this can lead to numerical instabilities, problems of this kind are challenging and require some care.
To lessen the impact of this, we split the pressure $\pressure = \backgroundPressure + \pressure'$ into a sum of the background pressure $\backgroundPressure(z)$ and perturbation $\pressure'(\bm{x}, t)$.
We split the density $\Qrho = \backgroundRho + \Qrho'$ in the same manner and arrive at
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence{ \left(
    \Qv \otimes \Qj + \bm{I} \pressure'
    \right)
  } + \viscFlux_\Qj
  =
  -g \bm{k} \Qrho'.
\end{equation}
Note that a similar and more complex splitting is performed in~\cite{muller2010adaptive,giraldo2008study}.
In contrast to this, we use the true compressible Navier-Stokes equations with minimal modifications.
\todo{Mention differences in splitting methods}
\section{Numerics}

\todo{LR: Hier würde ich noch ein paar einleitende Worte hinzufügen}
\todo{LK: CFL? Riemannsolver (no gradients)}
We use an \aderdg{}-scheme and a \muscl{} finite volume method.
Both can be considered as instances of the more general \textsc{PnPm} schemes of~\cite{dumbser2008unified}.
We use a Rusanov-style flux that is adapted to \pde{}s with viscous terms~\cite{gassner2008discontinuous,fambri2017space}.
The finite volume scheme is stabilized with the van Albada limiter\todo{LK: Which one is best? Cite vanalbada}.

The user can state dynamic \amr{} rules by supplying custom criteria that are evaluated point-wise.
We use an element local error estimate based on the total variation of the numerical solution. 
We use the fact that the total variation of a numerical solution is a perfect indicator for edges of a wavefront.
Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a sufficiently smooth function that maps the discrete solution at a point $\bm{x}$ to an arbitrary indicator variable.
The total variation (\textsc{tv}) of this function is defined by
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}
for each cell.
The operator $\Vert \cdot \Vert_1$ denotes the discrete $l_1$ norm in this equation.
We compute the integral efficiently with Gaussian quadrature over the collocated quadrature points.
%\todo{LK: If using limiting, add sentence about tv computation for fv}
\newcommand{\mean}{\mu}%
\newcommand{\std}{\sigma}%
\newcommand{\variance}{\std^2}%
\newcommand{\Trefine}{T_\text{refine}}%
\newcommand{\Tdelete}{T_\text{erase}}%
How can we decide whether a cell is important or not?
To resolve this conundrum, we compute the mean and the population standard deviation of the total variation of all cells.
It is important to note that we use the method of~\cite{chan1982updating} to compute the modes in a parallel and numerical stable manner.
A cell is then considered to contain significant information if its deviates from the mean more than a given threshold.
This criterion can be described formally by
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + \Trefine \sigma, \\
    \text{erase} & \text{if } \tv(\Q) < \mu + \Tdelete \sigma, \\
    \text{keep} & \text{otherwise}.
    \end{cases}
\end{equation}
The parameters $\Trefine > \Tdelete$ can be chosen freely.
Chebyshev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$ guarantees that we neither mark all cells for refinement nor for erasement.
This inequality holds for arbitrary distributions under the weak assumption that they have a finite mean $\mu$ and a finite standard deviation $\sigma$~\cite{wasserman2004all}.
Note that subcells are coarsened only if all subcells belonging to the coarse cell are marked for erasion.
In contrast to already published criteria which are either designed solely for the simulation of clouds~\cite{muller2010adaptive} or computationally expensive~\cite{fambri2017space}, our criterion works for arbitrary \pde{}s and yet, is easy to compute and intuitive.

We use a \muscl{} based finite volume limiting as proposed in~\cite{dumbser2016simple}.
Briefly, we first compute a candidate solution with our \aderdg{}-scheme and check whether it is valid.
A solution is considered to be invalid if it is a non-finite floating point number or if it is unphysical, i.e.\ if the pressure or density are not positive.
%\todo{LK: Mention dmp if using, mention tv-indicator for limiting if using}
In case of an invalid solution, we recompute the timestep with a our \muscl{} finite volume method.\todo{Source for the fv method}

@@

For the \textsc{cfd} scenarios, we use the following constants of fluids:
\begin{equation}
  \gamma = 1.4, \quad \Pr = 0.7, \quad c_v = 1.0.
\end{equation}

\todo{LK: Constants of fluids?}
A simple scenario is the Taylor-Green vortex.
It has an analytical solution in the incompressible limit which is given by
\begin{align}
  \label{eq:taylor-green}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) \cos(y) \\
- \cos(x) \sin(y) 
    \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \nicefrac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C.
  \end{split}
\end{align}
The constant $C = \nicefrac{100}{\gamma}$ governs the speed of sound and thus the Mach number $\text{Ma} = 0.1$~\cite{dumbser2016high}, the viscosity is set to $\mu = 0.1$.
We use a domain of size $(\SI{2 \pi}{\m} \times \SI{2 \pi}{\m})$ and impose the analytical solution at the boundary.
A comparison at time $t = 1.0$ of the analytical solution for the pressure with our approximation (figure xxx) shows an excellent agreement.\todo{fig, check time, fix order/cells.}
We used an \aderdg{}-scheme of order $5$ with a grid of $25^2$ cells.

The Arnold-Beltrami-Childress (\textsc{abc}) flow is similar to the Taylor-Green vortex but posses an analyitcal solution fir three-dimensions as well~\cite{tavelli2016staggered}.
It is defined in the domain \( \left[ -\SI{\pi}{\m}, \SI{\pi}{\m} \right]^3 \) as
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C.
  \end{split}
\end{align}
again under the assumption of incompressibility.
The constant $C = \nicefrac{100}{\gamma}$ is chosen as before.
We use a viscosity of $\mu = 0.01$ and analytical boundary conditions.
Figure xxx show a good agreement between the analytical solution and our approximation with an \aderdg{}-scheme of order $2$ with a mesh consisting of $27^3$ cells at time $t = \SI{0.1}{\s}$\todo{Figure}.

As a final example of standard flow scenarios, we consider the lid-driven cavity flow where the fluid is initially at rest, with $\Qrho = 1$ and $ \pressure(\bm{x}) = \nicefrac{100}{\gamma}$.
We consider a domain of size $\SI{1}{m} \times \SI{1}{\m}$ which is surrounded by no-slip walls.
The flow is driven entirely by the upper wall which has a velocity of $v_x = \SI{1}{\m/\s}$.
The simulation runs for $\SI{10}{\s}$
Again, our results have an excellent agreement with the reference solution of~\cite{ghia1982high}.
We used an \aderdg{}-method of order $3$ with a mesh of size $27^2$\todo{check!}.

Our stratified flow scenarios are described in terms of the potential energy
\begin{equation}
  \potT = T \left( \frac{p_0}{p} \right)^{R/c_p},
\end{equation}
with reference pressure $p_o$~\cite{muller2010adaptive,giraldo2008study}.
We use the constants
\begin{equation}\label{eq:atmosphere-constants}
    \gamma = 1.4 ,\qquad \Pr =  0.71 ,\qquad R = 287.058 ,\qquad p_0 = \SI{10e5}{\Pa}, \qquad g = \SI{9.8}{m/s^2},
\end{equation}
for all remaining scenarios.

We compute the initial background density and pressure by inserting the assumption of a constant background energy in \cref{eq:hydrostatic-balance}.
The background atmosphere is then perturbed.
Let
\begin{equation}
  \label{eq:radius}
  r^2 = \Vert \bm{x} - \bm{x_c} \Vert_2 %(x - x_0)^2 + (z - z_0)^2
\end{equation}
denote the euclidian distance between the spatial positions $\bm{x} = x, z$ and the center of a bubble $\bm{x_c} = x_c, z_c$.
In case of three-dimensional scenarios the vectors $\bm{x}$ and $\bm{x_c}$ also contain a y coordinate.
For our first scenario, the cosine bubble, we use a pertubation of the form
\begin{align}
  \label{eq:cos-pertubation}
  \pertubationPotT &= \begin{cases}
    \nicefrac{A}{2} \left[ 1 + \cos(\pi r) \right] & r \leq a, \\
    0 & r > a,
    \end{cases}
\end{align}
where $A$ denotes the maximal perturbation and $a$ is the size of the bubble.
We use the constants\todo{3d bubble is higher}
\begin{equation}\label{eq:cosine-bubble}
  A = \SI{0.5}{\K} \quad a = \SI{250}{\m} \quad x_c = \SI{500}{\m} \quad z_c = \SI{350}{\m}.
\end{equation}
For the three-dimensional bubble, we set $y_c = x_c = \SI{500}{\m}$ and lower the bubble to $z_c = \SI{260}{\m}$
We use a constant viscosity of $\mu = 0.1$ and a simulation time of $\SI{400}{\s}$ for 3d and $\SI{600}{\s}$ for 2d.

>>>>

\usepackage{caption}
\usepackage{subcaption}
\usepackage{xparse} % for NewDocumentCommand
\usepackage{etoolbox} % for notblank (brackets only when argument)
\usepackage{xstring} % for \IfSubStr
\usepackage{xpatch}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools} % for \mathclap
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage{varioref}
\usepackage{nicefrac}
\usepackage{physics} % for derivatives
\usepackage[separate-uncertainty]{siunitx}
\usepackage{hyperref}
\usepackage[noabbrev]{cleveref}
\newcommand{\creflastconjunction}{, and\nobreakspace} % use Oxford comma
\usepackage{todonotes}
\usepackage{multimedia}
\graphicspath{{../figures/}}
\let\boundary\undefined%
\crefformat{equation}{\eqA{}eq.~\eqB #2#1#3)}
% TODO: Fix for more than two equations!
\crefmultiformat{equation}{eqs.~\eqMultiA#2#1#3\eqMultiB}%
{ and \eqMultiA#2#1#3\eqMultiB}{, \eqMultiA#2#1#3\eqMultiB}{ and~\eqMultiA#2#1#3\eqMultiB}
\newcommand{\eqA}{}
\newcommand{\eqB}{(}
\newcommand{\eqMultiA}{(}
\newcommand{\eqMultiB}{)}
\DeclareRobustCommand{\pcrefSingle}[1]{%
\begingroup%
  \renewcommand{\eqA}{(}\renewcommand{\eqB}{}%
\cref{#1}%
\endgroup%
}
\DeclareRobustCommand{\pcrefMulti}[1]{%
\begingroup%
    \renewcommand{\eqMultiA}{}\renewcommand{\eqMultiB}{}%
    (\cref{#1})%
\endgroup%
}
\DeclareRobustCommand{\pcref}[1]{%
\IfSubStr{#1}{,}{\pcrefMulti{#1}}{\pcrefSingle{#1}}%
}

@@

We follow a different approach in this paper and use the \aderdg{} method of~\cite{dumbser2008unified} to simulate various computational fluid dynamics (\textsc{cfd}) scenarios with a focus on flows with stratified flows.
Due to the viscous components of the Navier-Stokes equations, it is not straightforward to apply the \dg{} method.
For high-order \dg{} methods, viscosity not only models the physical reality of the problem but also helps us by smoothing out oscillations and discontinuities, thus stabilizing the simulation.
To include viscosity, we use the numerical flux for the compressible Navier-Stokes equations of \citeauthor{gassner2008discontinuous}~\cite{gassner2008discontinuous}.
This flux has already been applied to the \aderdg{} method in~\cite{dumbser2010arbitrary}.
In contrast to this paper, we focus on the simulation of complex flows with a gravitational source term and a realistic background atmosphere.
Additionally, we use adaptive mesh refinement (\textsc{amr}) to increase the spatial resolution in areas of interest.
This has been shown to work well for the simulation of cloud dynamics~\cite{muller2010adaptive}.
%Insert references to Giraldo SE/DG bubbles~\cite{giraldo2008study}, 3D bubbles~\cite{giraldo2013implicit,kelly2012continuous}, bubble amr (müller~\cite{muller2010adaptive}).

@@

We make the following contributions:%
\begin{itemize}%
\item We extend the \exahypeengine{} to allow viscous terms following the aforementioned numerical scheme.
\item We use this to provide an implementation of the compressible Navier-Stokes equation.
  In addition, we tailor the equation set to stratified flows with gravitational source term.
  We want to emphasize here that we use a standard formulation of the Navier-Stokes equations as seen in the field of computational fluid mechanics and only use small modifcations of the governing equations, in contrast to a equation set that is tailored exactly to the application area.
\item We present a general \textsc{amr}-criterion that is based on the detection of outlier cells w.r.t.\ their total variation.
  Furthermore, we show how to utilize this criterion for stratified flows.
\item We evaluate our implementation with standard \textsc{cfd} scenarios and atmospheric flows and inspect the effectiveness of our proposed \amr{}-criterion.
  We thus inspect, whether our proposed general implementation can achieve results that are competitive with the state-of-the-art models that rely on heavily specified equations and numerics.
\end{itemize}
%
\section{Equation Set}
\newcommand{\diffCoeff}{\varepsilon}%
\newcommand{\hyperFluxDef}{
  \begin{pmatrix}
    \Qj \\
    \Qv  \otimes \Qj + \bm{I} \pressure  \\
    \Qv \cdot (\bm{I} \QE + \bm{I} \pressure)
  \end{pmatrix}
}%
\newcommand{\viscFluxDef}{
  \begin{pmatrix}
    0\\
     \stressT (\Q, \gradQ)  \\
     \Qv \cdot \stressT (\Q, \gradQ) - \kappa \gradient{T}
   \end{pmatrix}
}%
% \begin{equation}
%   \label{eq:ns-short}
%  \pdv{}{t} \bm{Q} + \divergence{\bm{F}(\bm{Q}, \gradient{\bm{Q}})} = \bm{S}(\bm{Q}) 
% \end{equation}
The compressible Navier-Stokes equations in the conservative form are defined as% \pcref{eq:conservation-law-gradient}
\begin{equation}
 \label{eq:equation-set} 
%  \begin{array}{l}
%  \text{mass cons.} \\
%  \text{momentum cons.} \\
%  \text{energy cons.} \\
%  \text{cont.\ gas} 
% \end{array}
% :
\quad
  \pdv{}{t}
  \underbrace{
  \begin{pmatrix}
    \Qrho\\
    \Qj\\
    \QE
    \end{pmatrix}}_{\Q}
  +
  \divergence{
  \underbrace{
  \left(
   \underbrace{\hyperFluxDef}_{\hyperFlux(\Q)}
+
\underbrace{\viscFluxDef}_{\viscFlux(\Q, \gradQ)}
  \right)}_{\flux(\Q, \gradQ)}}
 =
  \underbrace{
  \begin{pmatrix}
    \source[\Qrho\phantom{\Qrho}]\\
    \source[\Qj]\\
    \source[\QE]
    \end{pmatrix}}_{\source(\Q, \bm{x}, t)}
\end{equation}
with vector of conserved quantities $\Q$, flux $\flux(\Q, \gradQ)$ and source $\source(\Q)$.
Note that the flux can be split into a hyperbolic part $\hyperFlux(\Q)$,
which is identical to the flux of the Euler equations,
and a viscous part $\viscFlux(\Q, \gradQ)$.
The conserved quantities
\(\Q\)
% \begin{equation} 
%   \label{eq:conserved-variables}
%  \Q = \left( \Qrho, \Qj, \QE, \QZ \right),
% \end{equation}
are the density $\Qrho$, the two or three-dimensional momentum $\Qj$ and the energy density $\QE$.
The rows of \cref{eq:equation-set} are the conservation of mass, the conservation of momentum and the conservation of energy.

@@

Many realistic atmospheric flows can be described by a perturbation over a background state that is in hydrostatic equilibrium\todo{Mention source}
\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
i.e.\ a state, where the pressure gradient is exactly in balance with the gravitational source term $\source[\Qj] = - \bm{k} \Qrho g$.
The vector $\bm{k}$ is the unit vector pointing in $z$-direction.
The momentum equation is dominated by the background flow in this case.
Because this can lead to numerical instabilities, problems of this kind are challenging and require some care.
To lessen the impact of this, we split the pressure $\pressure = \backgroundPressure + \pressure'$ into a sum of the background pressure $\backgroundPressure(z)$ and perturbation $\pressure'(\bm{x}, t)$.
We split the density $\Qrho = \backgroundRho + \Qrho'$ in the same manner and arrive at
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence{ \left(
    \Qv \otimes \Qj + \bm{I} \pressure'
    \right)
  } + \viscFlux_\Qj
  =
  -g \bm{k} \Qrho'.
\end{equation}
Note that a similar and more complex splitting is performed in~\cite{muller2010adaptive,giraldo2008study}.
In contrast to this, we use the true compressible Navier-Stokes equations with minimal modifications.
\section{Numerics}
\todo{LR: Hier würde ich noch ein paar einleitende Worte hinzufügen}
We use an \aderdg{}-scheme and a \muscl{} finite volume method.
Both can be considered as instances of the more general \textsc{PnPm} schemes of~\cite{dumbser2008unified}.
We use a Rusanov-style flux that is adapted to \pde{}s with viscous terms~\cite{gassner2008discontinuous,fambri2017space}.
The finite volume scheme is stabilized with the van Albada limiter\todo{Cite vanalbada}.

The user can state dynamic \amr{} rules by supplying custom criteria that are evaluated point-wise.
We use an element local error estimate based on the total variation of the numerical solution. 
We use the fact that the total variation of a numerical solution is a perfect indicator for edges of a wavefront.
Let $\bm{f}(\bm{x}): \mathbb{R}^{N_\text{vars}} \to \mathbb{R}$ be a sufficiently smooth function that maps the discrete solution at a point $\bm{x}$ to an arbitrary indicator variable.
The total variation (\textsc{tv}) of this function is defined by
\begin{equation}
  \label{eq:tv}
  \tv \left[ f(\bm{x}) \right] =
  \Vert
\intdcell{ \vert \gradient{f \left( \bm{x} \right)} \vert }
\Vert_1
\end{equation}
for each cell.
The operator $\Vert \cdot \Vert_1$ denotes the discrete $l_1$ norm in this equation.
We compute the integral efficiently with Gaussian quadrature over the collocated quadrature points.
\newcommand{\mean}{\mu}%
\newcommand{\std}{\sigma}%
\newcommand{\variance}{\std^2}%
\newcommand{\Trefine}{T_\text{refine}}%
\newcommand{\Tdelete}{T_\text{erase}}%
How can we decide whether a cell is important or not?
To resolve this conundrum, we compute the mean and the population standard deviation of the total variation of all cells.
It is important to note that we use the method of~\cite{chan1982updating} to compute the modes in a parallel and numerical stable manner.
A cell is then considered to contain significant information if its deviates from the mean more than a given threshold.
This criterion can be described formally by
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + \Trefine \sigma, \\
    \text{erase} & \text{if } \tv(\Q) < \mu + \Tdelete \sigma, \\
    \text{keep} & \text{otherwise}.
    \end{cases}
\end{equation}
The parameters $\Trefine > \Tdelete$ can be chosen freely\todo{LK: Do we need the ineq.?}.
Chebyshev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$ guarantees that we neither mark all cells for refinement nor for erasement.
This inequality holds for arbitrary distributions under the weak assumption that they have a finite mean $\mu$ and a finite standard deviation $\sigma$~\cite{wasserman2004all}.
Note that subcells are coarsened only if all subcells belonging to the coarse cell are marked for erasion.
In contrast to already published criteria which are either designed solely for the simulation of clouds~\cite{muller2010adaptive} or computationally expensive~\cite{fambri2017space}, our criterion works for arbitrary \pde{}s and yet, is easy to compute and intuitive.

@@

For the standard \textsc{cfd} testing scenarios, we use the following constants of fluids:
\begin{equation}
  \gamma = 1.4, \quad \Pr = 0.7, \quad c_v = 1.0.
\end{equation}
A simple scenario is the Taylor-Green vortex.
It has an analytical solution in the incompressible limit which is given by
\begin{align}
  \label{eq:taylor-green}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \exp(-2 \mu t)
  \begin{pmatrix}
    \phantom{-}\sin(x) \cos(y) \\
- \cos(x) \sin(y) 
    \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= \exp(-4 \mu t) \, \nicefrac{1}{4} \left( \cos(2x) + \cos(2y) \right) + C.
  \end{split}
\end{align}
The constant $C = \nicefrac{100}{\gamma}$ governs the speed of sound and thus the Mach number $\text{Ma} = 0.1$~\cite{dumbser2016high}, the viscosity is set to $\mu = 0.1$.
We use a domain of size $(\SI{2 \pi}{\m} \times \SI{2 \pi}{\m})$ and impose the analytical solution at the boundary.
A comparison at time $t = 10.0$ of the analytical solution for the pressure with our approximation (figure xxx) shows an excellent agreement.\todo{fig, check time, fix order/cells.}
We used an \aderdg{}-scheme of order $5$ with a grid of $25^2$ cells.

\begin{figure}[htb]
  \centering
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics{paper_lid_driven_cavity}
    \caption{\label{fig:cavity-flow}Our approximation (lines) of the lid-driven cavity flow vs.\ reference solution (markers) of~\cite{ghia1982high}.}
  \end{subfigure}~%
  \begin{subfigure}[t]{0.5\textwidth}
    \centering
    \includegraphics{paper_abc_flow_velocity}
    \caption{\label{fig:abc-flow}\textsc{abc}-flow. We show the velocity in $x$-direction for slices through the domain, where all but the mentioned coordinates are held constant at a value of zero.
    The analytical solution is represented by lines, the numerical results by markers.}
  \end{subfigure}
  \caption{\textsc{cfd} scenarios}
  \label{fig:cdf-results}
\end{figure}

The Arnold-Beltrami-Childress (\textsc{abc}) flow is similar to the Taylor-Green vortex but posesses an analytical solution for three-dimensions as well~\cite{tavelli2016staggered}.
It is defined in the domain \( \left[ -\SI{\pi}{\m}, \SI{\pi}{\m} \right]^3 \) as
\begin{align}
  \label{eq:abc-flow}
  \begin{split}
  \Qrho(\bm{x}, t) &= 1,\\
  \Qv(\bm{x}, t) &= \phantom{-} \exp(-1\mu t)
  \begin{pmatrix}
    \sin(z) + \cos(y)\\
    \sin(x) + \cos(z)\\
    \sin(y) + \cos(x)
  \end{pmatrix}, \\
  \pressure(\bm{x}, t) &= -\exp(-2 \mu t) \, \left(\cos(x)\sin(y) + \sin(x)\cos(z) + \sin(z)\cos(y)\right)
  + C.
  \end{split}
\end{align}
again under the assumption of incompressibility.
The constant $C = \nicefrac{100}{\gamma}$ is chosen as before.
We use a viscosity of $\mu = 0.01$ and analytical boundary conditions.
Our results (\cref{fig:abc-flow}) show a good agreement between the analytical solution and our approximation with an \aderdg{}-scheme of order $3$ with a mesh consisting of $27^3$ cells at time $t = \SI{0.1}{\s}$.

As a final example of standard flow scenarios, we consider the lid-driven cavity flow where the fluid is initially at rest, with $\Qrho = 1$ and $ \pressure(\bm{x}) = \nicefrac{100}{\gamma}$.
We consider a domain of size $\SI{1}{m} \times \SI{1}{\m}$ which is surrounded by no-slip walls.
The flow is driven entirely by the upper wall which has a velocity of $v_x = \SI{1}{\m/\s}$.
The simulation runs for $\SI{10}{\s}$
Again, our results (\cref{fig:cavity-flow}) have an excellent agreement with the reference solution of~\cite{ghia1982high}.
We used an \aderdg{}-method of order $3$ with a mesh of size $27^2$.

Our stratified flow scenarios are described in terms of the potential temperature
\begin{equation}
  \potT = T \left( \frac{p_0}{p} \right)^{R/c_p},
\end{equation}
with reference pressure $p_o$~\cite{muller2010adaptive,giraldo2008study}.
We use the constants
\begin{equation}\label{eq:atmosphere-constants}
    \gamma = 1.4 ,\qquad \Pr =  0.71 ,\qquad R = 287.058 ,\qquad p_0 = \SI{10e5}{\Pa}, \qquad g = \SI{9.8}{m/s^2},
\end{equation}
for all remaining scenarios.

We compute the initial background density and pressure by inserting the assumption of a constant background energy in \cref{eq:hydrostatic-balance}.
The background atmosphere is then perturbed.
We set the density and energy at the boundary such that it corresponds to the background atmosphere.
Furthermore, to ensure that the atmosphere stays in hydrostatic balance, we need to impose the viscous heat flux
\begin{equation}
  \label{eq:atmosphere-bc}
  \viscFlux_{\QE} = \kappa \pdv{\overline{T}}{z}.% =
%\kappa \frac{R \overline{\theta} \left(\frac{\overline{p}{\left (z \right )}}{p_{0}}\right)^{\frac{R}{c_{p}}} \frac{d}{d z} \overline{p}{\left (z \right )}}{c_{p} \overline{p}{\left (z \right )}}.
\end{equation}
at the boundary.
In this equation, $\overline{T}(z)$ is the background temperature at position $z$, which can be computed from \cref{eq:hydrostatic-balance,eq:eos}~\cite{giraldo2008study}.
